Get utilization standards and piece size by FMU and township

### 2021-09-29 analysis/InverseTaperModels.Rmd
almost there

```{r libraries, echo=FALSE, message=FALSE}
#setwd("g:/Shared drives/FES_Data_Sharing (Forestry)/Manuscripts/Historical Residue Supply/Analysis")
library(tidyverse)
library(ggthemes)
library(gridExtra)
library(viridis)
library(skimr)
library(scales)
library(DescTools)
library(zoo)
```

@@

diameters <- c(15,15.5,16,16.5,17,17.5,18,18.5,19,19.5,20,20.5,21,21.5,22,22.5,23,23.5,24,24.5,25,25.5,26,26.5,27,27.5,28,28.5,29,29.5,30
)

# lodgepole pine example

b0 <- 11.8367
b1 <- 0.246
b2 <- 2.1264
db0 <- 0.2569
db1 <- 1.02741
stumpht <- 0.15
stumpd <- 15
topD <- 11
minloglength <- 4.88

piece_size <- 0.2895

dguess <- 20.6
dbh <- dguess
dbhib <- (dbh - db0)/db1

height = function(dbh,b0,b1){
        ht <- b0 * dbh^b1
        return(ht)
}

ht <- height(dbh,b0,b1)
dbhib
ht

taper = function(h,dbh,b0,b1,b2){
  ht <- height(dbh,b0,b1)
  num <- ht - h
  den <- ht - 1.3
  d <- sqrt(dbh^2 * (h/1.3)^(2-b2) * (num/den))
  return(d)
}

taperTop = function(h,dbh,topD,b0,b1,b2){
        ht <- height(dbh,b0,b1)
        num <- ht - h
        den <- ht - 1.3
        toast <- sqrt(dbh^2 * (h/1.3)^(2-b2) * (num/den)) - topD
        return(toast)
      }

ht <- height(dbh,b0,b1)

merchht <- uniroot(taperTop, dbh=dbh,topD=topD,b0=b0, b1=b1,  b2=b2,lower=0,upper=ht)$root
loglength <- merchht - stumpht

      obArea = function(h){
        r <- taper(h,dbh,b0,b1,b2)
        r <- r/200
        area <- pi * r^2
        return(area)
      }

      ibArea = function(h){
        r <- taper(h,dbh,b0,b1,b2)
        #  r <- max(0,(r - db0)/db1)
        r <- (r - db0)/db1
        r <- max(0,r/200)
        area <- pi * r^2
        return(area)
      }

      obVolume <-
        integrate(obArea,lower=stumpht,upper=merchht)

      totalOBVolume <-
        integrate(obArea,lower=0,upper=ht)

      totalIBVolume <-
        integrate(Vectorize(ibArea),lower=0,upper=ht)

@@

      obVol <- if(loglength < minloglength){
        0
      } else {
        obVolume$value
      }

      ibVolume <-
        integrate(Vectorize(ibArea),lower=stumpht,upper=merchht)

      ibVol <- if(loglength < minloglength){
        0
      } else {
        ibVolume$value
      }

      merchVol <- ibVol
      stemBarkVolume <- obVol - ibVol
      topVolume <-
        integrate(Vectorize(ibArea),merchht,upper=ht)
      topObVol <- integrate(Vectorize(obArea),merchht,upper=ht)
      topBarkVolume <- topObVol$value - topVolume$value

      topVol <- topVolume$value
      top2merchRatio <- topVol/ibVol

      topBarkRatio <- topBarkVolume/totalBarkVolume
      totalIBVolume
      topVolRatio <- topVolume$value/totalIBVolume$value

      topVolume

merchht
loglength
obVol
ibVol
top2merchRatio

# ecozones <- c(4,5,6,9,10,14)
# speciesList <- c(",ABIE,LAS,,",",ABIE,BAL,,",",BETU,PAP,PAP,",",LARI,LAR,,",",PICE,ENG,,",",PICE,GLA,,",",PICE,MAR,,",
#                  ",PINU,BAN,,",",PINU,CON,LAT,",",POPU,TRE,,",",POPU,BAL,,",",PSEU,MEN,GLA,")
# species2 <- c("Fa","Fb","Bw","Lt","Se","Sw","Sb",
#               "Pj","Pl","Aw","Pb","Fd")
# header <- "Tree_num,Province,Ecozone,Genus,Species,Variety,DBH,Height"
# header2 <- "Tree_num,Province,Ecozone,Genus,Species,Variety,DBH,Height,merchVol,topVolRatio,topBarkRatio"
# 
# cat(sprintf("%s\n",header),file = "../data/bioCalcInput.csv", append = FALSE)
# cat(sprintf("%s\n",header2),file = "../data/ratios.csv", append = FALSE)
# 
# taperCoefficients <- read.csv(file = '../data/taperCoefficients.csv')
# trimAllowance <- 0.05 # m  or 5 cm
# minPieceLength <- 2.44 + trimAllowance
# 
# treenum <- 0
# spcnt <- 0
# for (sp in species2) {
#   spcnt <- spcnt + 1
#   t <- taperCoefficients %>% filter(species == species2[[spcnt]])
#   b0 <- t$b0
#   b1 <- t$b1
#   b2 <- t$b2
#   topD <- t$uTopD
#   stumpht <- t$uStumpHt
#   minloglength <- t$uMinlen
#   
#   db0 <- t$db0
#   db1 <- t$db1
#   
#   for (ez in ecozones){
#     for (val in diameters){
#       dbh <- val
#       dbhib <- (dbh - db0)/db1
#       
#       height = function(dbh,b0,b1){
#         ht <- b0 * dbh^b1
#         return(ht)
#       }
#       
#       
#       
#       taper = function(h,dbh,b0,b1,b2){
#         ht <- height(dbh,b0,b1)
#         num <- ht - h
#         den <- ht - 1.3
#         d <- sqrt(dbh^2 * (h/1.3)^(2-b2) * (num/den))
#         return(d)
#       }
#       
#       taperTop = function(h,dbh,topD,b0,b1,b2){
#         ht <- height(dbh,b0,b1)
#         num <- ht - h
#         den <- ht - 1.3
#         toast <- sqrt(dbh^2 * (h/1.3)^(2-b2) * (num/den)) - topD
#         return(toast)
#       }
#       
#       ht <- height(dbh,b0,b1)
#       
#       merchht <- uniroot(taperTop, dbh=dbh,topD=topD,b0=b0, b1=b1, b2=b2,lower=0,upper=ht)$root
#       loglength <- merchht - stumpht
#       
#       # merchht <- if(sp == "Aw"){
#       #   merchht <- 15
#       # } else {
#       #   merchht <- merchht
#       # }
#       # 
#       
#       
#       
#       
#       obArea = function(h){
#         r <- taper(h,dbh,b0,b1,b2) 
#         r <- r/200
#         area <- pi * r^2
#         return(area)
#       }
#       
#       ibArea = function(h){
#         r <- taper(h,dbh,b0,b1,b2) 
#         #  r <- max(0,(r - db0)/db1)
#         r <- (r - db0)/db1
#         r <- max(0,r/200)
#         area <- pi * r^2
#         return(area)
#       }
#       
#       obVolume <-
#         integrate(obArea,lower=stumpht,upper=merchht)
#       
#       totalOBVolume <-
#         integrate(obArea,lower=0,upper=ht)
#       
#       totalIBVolume <-
#         integrate(Vectorize(ibArea),lower=0,upper=ht)
#       
#       totalBarkVolume <- totalOBVolume$value - totalIBVolume$value
#       
#       obVol <- if(loglength < t$uMinlen){
#         0
#       } else {
#         obVolume$value
#       }
#       
#       ibVolume <- 
#         integrate(Vectorize(ibArea),lower=stumpht,upper=merchht)
#       
#       
#       
#       ibVol <- if(loglength < t$uMinlen){
#         0
#       } else {
#         ibVolume$value
#       }
#       
#       merchVol <- ibVol
#       stemBarkVolume <- obVol - ibVol
#       topVolume <-  
#         integrate(Vectorize(ibArea),merchht,upper=ht)
#       topObVol <- integrate(Vectorize(obArea),merchht,upper=ht)
#       topBarkVolume <- topObVol$value - topVolume$value 
#       
#       topVol <- topVolume$value
#       top2merchRatio <- topVol/ibVol
#     
#       
#       topBarkRatio <- topBarkVolume/totalBarkVolume
#       totalIBVolume
#       topVolRatio <- topVolume$value/totalIBVolume$value
#       
#       topVolume
#       top2merchRatio
#       logs <- loglength/(4.88 + 0.05)
#       toploglength <- ((logs - floor(logs)) * (4.88 + 0.05))
#       logs
#       toploglength
#       logs - floor(logs)
#       (logs - floor(logs)) * 4.93
#  #     oFile <- file("../data/output.csv")
#       treenum <- treenum + 1
#       stuff <- paste0(",AB,",ez,speciesList[[spcnt]])
#       
#       cat(sprintf("%i%s%6.2f%s%6.2f\n",treenum,stuff,dbh,",",ht),
#           file = "../data/bioCalcInput.csv",append = TRUE)
#       cat(sprintf("%i%s%6.2f%s%6.2f%s%f%s%f%s%f\n",treenum,stuff,dbh,",",ht,",",
#                   merchVol,",",topVolRatio,",",topBarkRatio),
#           file = "../data/ratios.csv",append = TRUE)
#     }
#   } 
#}
#closeAllConnections()

>>>>

```{r libraries, echo=FALSE, message=FALSE}
setwd("g:/Shared drives/FES_Data_Sharing_Forestry/Manuscripts/histressupp/analysis")
library(tidyverse)
library(ggthemes)
library(gridExtra)
library(viridis)
library(skimr)
library(scales)
library(DescTools)
library(zoo)
```

@@

taperCoefficients <-
  read.csv(file = '../data/taperCoefficients.csv', fileEncoding = 'UTF-8-BOM')
taperCoefficients

load("../data/dbhCalcInput.RData")

dbhCalcInput <- dbhCalcInput %>%
  filter(!is.na(max_species))  %>%
  mutate(
    species = recode(
      max_species,
      `pice_gla` = "Sw",
      `pice_mar` = "Sb",
      `pice_eng` = "Se",
      `pinu_con` = "Pl",
      `pinu_ban` = "Pj",
      `pinu_spp` = "Pl",
      `popu_tre` = "Aw",
      `abie_bal` = "Fb",
      `abie_las` = "Fa",
      `lari_lar` = "Lt"
    )
  )

dbhCalcInput

t <- left_join(dbhCalcInput, taperCoefficients, by = "species")
t <- t %>%
  select(-max_species,-uStumpHt, -uStumpD, -uTopD, -uMinlen, -sppgrp)

outfile <- file("../data/output.txt")

for (irow in 1:nrow(t))
{
  b0 <- t[irow, ]$b0
  b1 <- t[irow, ]$b1
  b2 <- t[irow, ]$b2
  db0 <- t[irow, ]$db0
  db1 <- t[irow, ]$db1
  stumpht <- t[irow, ]$stumpht / 100
  stumpd <- t[irow, ]$stumpd
  topD <- t[irow, ]$topd
  minloglength <- t[irow, ]$merch_len
  piece_size <- t[irow, ]$piece_size

  mindbh <- 15
  maxdbh <- 30

  lower <- mindbh
  upper <- maxdbh
  guess <- 22.5
  tolerance <- 0.0005

  obArea = function(h) {
    r <- taper(h, dbh, b0, b1, b2)
    r <- r / 200
    area <- pi * r ^ 2
    return(area)
  }

  ibArea = function(h) {
    r <- taper(h, dbh, b0, b1, b2)
    #  r <- max(0,(r - db0)/db1)
    r <- (r - db0) / db1
    r <- max(0, r / 200)
    area <- pi * r ^ 2
    return(area)
  }

  count <- 0
  repeat {
    # do something
    count <- count + 1
    dbh <- guess
    dbhib <- (dbh - db0) / db1

    height = function(dbh, b0, b1) {
      ht <- b0 * dbh ^ b1
      return(ht)
    }

    ht <- height(dbh, b0, b1)
    dbhib
    ht

    taper = function(h, dbh, b0, b1, b2) {
      ht <- height(dbh, b0, b1)
      num <- ht - h
      den <- ht - 1.3
      d <- sqrt(dbh ^ 2 * (h / 1.3) ^ (2 - b2) * (num / den))
      return(d)
    }

    taperTop = function(h, dbh, topD, b0, b1, b2) {
      ht <- height(dbh, b0, b1)
      num <- ht - h
      den <- ht - 1.3
      toast <-
        sqrt(dbh ^ 2 * (h / 1.3) ^ (2 - b2) * (num / den)) - topD
      return(toast)
    }

    ht <- height(dbh, b0, b1)

    merchht <-
      uniroot(
        taperTop,
        dbh = dbh,
        topD = topD,
        b0 = b0,
        b1 = b1,
        b2 = b2,
        lower = 0,
        upper = ht
      )$root
    loglength <- merchht - stumpht

    obVolume <-
      integrate(obArea, lower = stumpht, upper = merchht)

    totalOBVolume <-
      integrate(obArea, lower = 0, upper = ht)

    totalIBVolume <-
      integrate(Vectorize(ibArea), lower = 0, upper = ht)

@@

    obVol <- if (loglength < minloglength) {
      0
    } else {
      obVolume$value
    }

    ibVolume <-
      integrate(Vectorize(ibArea), lower = stumpht, upper = merchht)

    ibVol <- if (loglength < minloglength) {
      0
    } else {
      ibVolume$value
    }

    merchVol <- ibVol
    stemBarkVolume <- obVol - ibVol
    topVolume <-
      integrate(Vectorize(ibArea), merchht, upper = ht)
    topObVol <- integrate(Vectorize(obArea), merchht, upper = ht)
    topBarkVolume <- topObVol$value - topVolume$value

    topVol <- topVolume$value
    top2merchRatio <- topVol / ibVol

    topBarkRatio <- topBarkVolume / totalBarkVolume
    # totalIBVolume
    topVolRatio <- topVolume$value / totalIBVolume$value

    # topVolume
    #
    # merchht
    # loglength
    # obVol
    # top2merchRatio
    # ibVol

    # exit if the condition is met
    if (abs(piece_size - ibVol) <= tolerance)
      break
    if (ibVol > piece_size) {
      upper <- guess
    } else {
      lower <- guess
    }
    guess <- (upper + lower) / 2

  }

  calculated_dbh <- guess
  writeLines(c(t[irow,]$species,count,calculated_dbh,ht),outfile)
}

close(outfile)

### 2021-09-29 analysis/get_util_piece_size.Rmd
almost there

```{r libraries, echo=FALSE, message=FALSE}
#setwd("/home/gwa/Work/historical-residue/Analysis")
setwd("f:/Shared Drives/FES_Data_Sharing_Forestry/Manuscripts/histressupp/analysis")

@@

save(dbhCalcInput, file="../data/dbhCalcInput.Rdata")

>>>>

```{r libraries, echo=FALSE, message=FALSE}
#setwd("/home/gwa/Work/historical-residue/Analysis")
setwd("g:/Shared Drives/FES_Data_Sharing_Forestry/Manuscripts/histressupp/analysis")

@@

save(dbhCalcInput, file="../data/dbhCalcInput.RData")
